Week 3 Notebook

### 2022-02-25 docs/3. SNA - Assignments/Week3-Assignment.Rmd
2.24.22
I think I fixed the network objects

Okay, this time I REALLY fixed them AND kinda understand the metadata

```
## Working from the example R syntax:
1.   calculate network density; 
2.   degree centrality, and 
3.   centralization measures for your network data.

## Write up your results briefly:
*   focus on interpretation of the results. 
*   Don’t list statistics or provide code, 
*  **the goal is to choose a few interesting results and provide an interpretation.**

## Briefly describe the dataset:
1.   identify initial network format; 
2.   describe and identify the nodes 
     including: 
        *  how many nodes are in the dataset; 
3.   what constitutes a tie or edge
     including:
        *  how many ties, 
        *   whether ties are directed/undirected,
        *   whether ties are weighted/binary, 
        *   and how to interpret the value of the tie if any.
        *   Whether or not there are edge attributes 
            that might be used to:
            *   subset data, or 
            *   stack multiple networks. 
Not every feature of the network needs to be described, but **description should orient reader to the network data and provide any necessary context for the results provided.**

## Provide at least two or three noteworthy results, including:
1.   the relevant statistics and interpretation. 
        For example, explaining which node(s) are most central and which are least central. 
2.   Discuss (with any related evidence) whether or not the node(s) behavior is in line with or violates expectations based on the degree centrality measure. 
3.   What do you make of network density and centralization measures?  
4.   What about differences in the binary 100 million threshold network vs valued trade flows network. 
 5.  Check tutorial for other ideas of results to discuss.

> Distill is a publication format for scientific and technical writing, native to the web.
>
> Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.

>>>>

```

#        *Write up your results briefly: *
*   focus on interpretation of the results. 
*   Don’t list statistics or provide code, 
*  **the goal is to choose a few interesting results and provide an interpretation.**

##        *Briefly describe the dataset:*
1.   identify initial network format; 
2.   describe and identify the nodes 
     including: 
        *  how many nodes are in the dataset; 
3.   what constitutes a tie or edge
     including:
        *  how many ties, 
        *   whether ties are directed/undirected,
        *   whether ties are weighted/binary, 
        *   and how to interpret the value of the tie if any.
        *   Whether or not there are edge attributes 
            that might be used to:
            *   subset data, or 
            *   stack multiple networks.

##        *Provide at least two or three noteworthy results*, including:
1.   the relevant statistics and interpretation. 
        For example, explaining which node(s) are most central and which are least central. 
2.   Discuss (with any related evidence) whether or not the node(s) behavior is in line with or violates expectations based on the degree centrality measure. 
3.   What do you make of network density and centralization measures?  
4.   What about differences in the binary 100 million threshold network vs valued trade flows network. 
 5.  Check tutorial for other ideas of results to discuss.

#  Working from the example R syntax:
1.   calculate network density; 
2.   degree centrality, and 
3.   centralization measures for your network data. 
# Import & Describe Network

# 1.   calculate network density
# 2.   degree centrality
# 3.   centralization measures for your network data.

# 
> Distill is a publication format for scientific and technical writing, native to the web.
>
> Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.

### 2022-02-25 docs/3. SNA - Assignments/sna_Week3/Week3_Notebook.Rmd
2.22.22 late on this past weekend's committs
2.24.22
I think I fixed the network objects

Okay, this time I REALLY fixed them AND kinda understand the metadata

---
title: "R Notebook"
output: html_notebook
---

```{r setup, Libraries, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

# Document Creation
library(knitr)
library(markdown)
library(yaml)
library(distill)
library(blogdown)

# Basic Analysis
library(ggplot2)
library(tidyverse)
library(clipr)
library(dbplyr)
library(data.table)
library(stringr)
library(stringi)
library(tibble)
library(tidyr)

# SNA - Social Network Analysis

library(tergm)
library(ergm)
library(network)
library(networkDynamic)
library(networkDynamicData)

library(statnet)
library(igraph)
library(igraphdata)
library(tsna)
library(sna)

```{r Set Working Directory, include=FALSE}
getwd()

setwd("C:/Users/white/iCloudDrive/Documents/1_DACSS_UMassAmherst/Dacss_RProjects/3. Sp22_rProjects/Dacss_679- Coursework/GIT-ajw_Learning-SocialNetworkAnalysis/docs/3. SNA - Assignments/sna_Week3")

## Basic R: Import & Clean Data

```{r Read raw data into R}

library(readr)

indx_SmugglerCove_recip <- read_csv("C:/Users/white/iCloudDrive/Documents/1_DACSS_UMassAmherst/Dacss_RProjects/3. Sp22_rProjects/Dacss_679- Coursework/GIT-ajw_Learning-SocialNetworkAnalysis/docs/3. SNA-Data/index-by_name-Smuggler_Cove-cocktail_recipies.csv", 
    col_names = FALSE, 
    col_types = cols(X1 = col_skip(),
                     X4 = col_skip()), 
                    na = "empty")

View(indx_SmugglerCove_recip)
summary(indx_SmugglerCove_recip)

 # Smuggler's Cove

smugCovIdx <- indx_SmugglerCove_recip

```{r Getting Information on a Dataset:, echo=FALSE}

# list the structure of smugCovIdx
str(smugCovIdx)

# class of an object (numeric, matrix, data frame, etc)
class(smugCovIdx)

# list the variables in smugCovIdx
names(smugCovIdx)

# dimensions of an object
dim(smugCovIdx)
```

```{r & Clean Data}

  library(stringr)
      # Drop NA rows
smugCovIdx <- na.omit(smugCovIdx)

# DATA Cleaning

#     Extract variables
#####

          # extract first word of the column, adds values in new Column at the end of DF

                    #1st pastIt
passIt <- smugCovIdx

colnames(passIt)
                    #2nd passIt
passIt$substring_first <- str_extract(passIt$X5,"(\\w+)")

#####

#####
#     Drop it like...
#####
      # DROP COLUMNS FROM DATA FRAME

                      #3rd PassIt
passIt = subset(passIt, 
                       select = -c(3)
                       )
colnames(passIt)
#####

#####
#     Give Variables a human name
#####

# rename column names with human language

                      # Return it
smugCovIdx <- passIt

names(smugCovIdx )[1] <- "dnkName"
names(smugCovIdx )[2] <- "dnkIng"
names(smugCovIdx )[3] <- "ingType"

colnames(smugCovIdx)
```

#   Project Varriables

    Raw Data:
  indx_SmugglerCove_recip

    DataFrame:
  smugCovIdx
            colname: "dnkName" "dnkIng"  "ingType"
  network_nodes <- smugCovIdx << "dnkName"

    Matrix:
  smug_matrix
  network_edgelist <- smug_matrix

    Network Objects: EDGE LIST
  STATNET: smugEL_sn
  IGRAPH: smugEL_ig

  Edge Vetex:
              dnkName
  Edge Attributes:
                  dnkIng
                  ingType

# Network Info
  Network Objects: EDGE LIST
                  STATNET: smugEL_sn
                  IGRAPH: smugEL_ig
                          ingTypeEL_Dig

```{r EDGELIST: UNDIRECTED Network Objects: statnet & igraph, include=FALSE}

#Lets make some network objects

colnames(smugCovIdx) 
# tell R data should be understood as a matrix
smug_matrix = as.matrix(smugCovIdx,
                        matrix.type="edgelist",
                        directed=FALSE) 
head.matrix(smug_matrix)

network_edgelist <- smug_matrix

  network_nodes <- smugCovIdx 
    network_nodes<- 
      select(network_nodes,
             dnkName)

        ### View(network_nodes)

######
      ######**igraph** 
######
        #igraph object 
network_ig <- 
  graph.edgelist(
    as.matrix(
      smugCovIdx[1:2]), 
    directed=F)
                ## create new edge attribute
  E(network_ig)$dnkIng<-smugCovIdx$dnkIng
  E(network_ig)$ingType<-smugCovIdx$ingType
  #create vertex attributes in igraph
 V(network_ig)[match(V(network_ig)$dnkName,network_nodes$dnkName)]$dnkName<-
   network_nodes$dnkName

      summary(edge_attr(network_ig))

  # head(network_ig)
  # summary(network_ig)
      ######

#####
      #####**statnet** 
#####    
  network_edgelist <- smugCovIdx

  network_nodes <- smugCovIdx 
  network_nodes<- 
    select(network_nodes,
           dnkName)
summary(network_nodes)
View(network_nodes)

  #Now let's create a statnet object
  network_statnet <- network(network_edgelist, 
                             matrix.type = "edgelist",
                             vertex.attr = network_nodes,
                             multiple = T)

  network_statnet%e%'dnkIng'  <-
    network_edgelist$dnkIng

  network_statnet%e%'ingType'<- 
    network_edgelist$ingType

       # list the names: vertex attributes
  view(network_statnet%v% "vertex.names")

summary(network_statnet)
        #####

 ## Network Objects: EDGE LIST
              ##**STATNET:**
  smugEL_sn <- network_statnet

                ##**IGRAPH:** 
  smugEL_ig <- network_ig

```{r EDGELIST: IGRAPH: DIRECTED Network: ingType, include=FALSE}

#Lets make some network objects

colnames(smugCovIdx) 
# tell R data should be understood as a matrix
smug_matrix = as.matrix(smugCovIdx,
                        matrix.type="edgelist",
                        directed=FALSE) 
head.matrix(smug_matrix)

network_edgelist <- smug_matrix

  nodes_ingType <- smugCovIdx 
    nodes_ingType<- 
      select(nodes_ingType,ingType)
view(nodes_ingType)
        ### View(nodes_ingType)

######
      ######**igraph** 
######
        #igraph object: ingType
d_network_ig <- 
  graph.edgelist(
    as.matrix(
      smugCovIdx[2:3]), 
    directed=T)

vertex.attributes(d_network_ig)
                ## create new edge attribute
  E(d_network_ig)$dnkIng<-smugCovIdx$dnkIng
  E(d_network_ig)$dnkName<-smugCovIdx$dnkName

  #create vertex attributes in igraph
 V(d_network_ig)[match(V(d_network_ig)$ingType,nodes_ingType$ingType)]$ingType<- nodes_ingType$ingType

   View(d_network_ig)
   head(d_network_ig)
  summary(d_network_ig)
  summary(edge_attr(d_network_ig))
  summary(vertex.attributes(d_network_ig))

   ingTypeEL_Dig <- d_network_ig
```

## Smuggler's Cove Index- NETWORK: EDGELIST: Features & Descriptions

@@

            # list the names:  edge attributes
  network::list.edge.attributes(smugEL_sn)
      #   summarize: numeric network attribute
  summary(smugEL_sn%v%"dnkName")

@@

smugEL_sn

[1] "network"

Network attributes:
  vertices = 274 
  directed = TRUE 
  hyper = FALSE 
  loops = FALSE 
  multiple = TRUE 
  bipartite = FALSE 
  total edges= 672 
    missing edges= 0 
    non-missing edges= 672

 Vertex attribute names: 
    dnkName vertex.names

 Edge attribute names: 
    dnkIng ingType

@@

gtrans(smugEL_sn)
proportion of transitive triads

symmetrize( smugEL_sn, rule = "weak") 
symmetrize so i<->j iff i->j OR i<-j

symmetrize( smugEL_sn, rule = "strong") 
symmetrize so i<->j iff i->j OR i<-j

dyad.census(smugEL_sn) 
MAN dyad census

triad.census(smugEL_sn) 
standard directed triad census

@@

    # Generate a list the names of vertex attributes
  V(smugEL_ig)$name

    # access edge attribute  names
  igraph::vertex_attr_names(smugEL_ig)

    # Generate a list the names of edge attributes  
  igraph::edge_attr_names(smugEL_ig)

## Network Attributes
 summary(smugEL_ig)
          # Vertices =  
          # Directed = 
          # hyper = 
          # loops = 
          # multiple = 
          # bipartite = 
          # total edges = 
                # missing edges =
                # non-missing edges = 
          # Vertex attribute names:
          # Edge Attribute Names: 
```
smugEL_ig

[1] "igraph"

Network attributes:
  vertices = 274 
  directed = FALSE
  hyper = FALSE 
  loops = FALSE 
  multiple = TRUE 
  bipartite = FALSE 
  total edges= 672 
    missing edges= 0 
    non-missing edges= 672

 Vertex attribute names: 
    dnkName vertex.names

 Edge attribute names: 
    dnkIng ingType

```{r IGRAPH: DIRECTED Network: ingType, echo=TRUE}

      # class of an object
    class(ingTypeEL_Dig)

@@

      # Count Vertices
    vcount(ingTypeEL_Dig)

      # Count Edges
    ecount(ingTypeEL_Dig)

## Network Mode
    # Bipartitie / Unimodal
    is_bipartite(ingTypeEL_Dig)

## Edges    
    # Un/directed 
    is_directed(ingTypeEL_Dig)

    # weighted or unweighted?
    is_weighted(ingTypeEL_Dig)

    # edge weights
    E(ingTypeEL_Dig)$weight

@@

    # Generate a list the names of vertex attributes
  V(ingTypeEL_Dig)$name

    # access edge attribute  names
  igraph::vertex_attr_names(ingTypeEL_Dig)

    # Generate a list the names of edge attributes  
  igraph::edge_attr_names(ingTypeEL_Dig)

## Network Attributes
 summary(ingTypeEL_Dig)
          # Vertices =  
          # Directed = 
          # hyper = 
          # loops = 
          # multiple = 
          # bipartite = 
          # total edges = 
                # missing edges =
                # non-missing edges = 
          # Vertex attribute names:
          # Edge Attribute Names:

@@

#     find average shortest path for network
    average.path.length(ingTypeEL_Dig,directed=T)

@@

#     isolating the distances between specific node
    all_shortest_paths(ingTypeEL_Dig,
                       "Strozzi","Tornabuoni", 
                       weights=NA)$res
## Note: manually tell igraph to ignore edge weights.
```

@@

#     Calculate distances between two nodes
    distances(ingTypeEL_Dig,"Bischeri","Ridolfi")

#     Calculate distance between two nodes using unweighted edges
    distances(ingTypeEL_Dig,"Bischeri", "Castellani",weights=NA)

@@

    # Number of components
 igraph::components(smugEL_ig)$no 
      ##**NOTE:** $no prevents a list of comonents from being printed

@@

```{r Statnet: , echo=TRUE}
      # Basic R: get summary statistics for node attributes
 summary(nodes_ingTypes)
      # provides range, mean, median and standard deviation of a vector.

@@

smug_attributes <- smugCovIdx
smug_edj <- smugCovIdx[2:3]
View(smug_edj)
#V1: dnkIng
#V2: ingType

@@

V(smugNet)$dnkIng <- 
  smug_attributes$dnkIng[match(V(smugNet)$name, smug_attributes$ingType)]

V(smugNet)$dnkName <- 
  smug_attributes$dnkName[
    match(V(
      smugNet)$name,
      smug_attributes$dnkIng)
    ]
summary(smugNet)

@@

V(smugNet)$ingType <- 
  smug_attributes$ingType[match(V(smugNet)$name, smug_attributes$ingType)]

V(smugNet)$dnkName <-
  smug_attributes$dnkName[match(V(smugNet)$name, smug_attributes$dnkName)]

@@

```{r iGraph: Undirected: DEGREE CENTRALITY: ingType, echo=TRUE}
# graph the SMUG network

smug_attributes <- smugCovIdx
smug_edj <- smugCovIdx[2:3]
View(smug_edj)
#V1: dnkIng
#V2: ingType

@@

V(smugNet)$dnkName <-
  smug_attributes$dnkName[match(V(smugNet)$name, smug_attributes$dnkIng)]

@@

 # differentiate nodes based on their attributes
 gplot(smugEL_sn, 
       displaylabels=TRUE,
       label.cex=.75,
       label.col="black",
       vertex.col=att$dnkName
 )

@@

graph.density(ingTypeEL_Dig)
    ## Density captures, at the macro-level, the proportion of dyads that are present over the possible total number of dyads in the network

@@

ingTypeEL_DigNod <- vcount(ingTypeEL_Dig)

ingTypeEL_DigDen <- graph.density(ingTypeEL_Dig)

@@

```{r iGraph: Directed: Degree Centrality}
# Degree centrality is calculated using the degree function in R. It returns how many edges each node has.
view(degree(ingTypeEL_Dig))

ingT_deg <- ingTypeEL_Dig
V(ingT_deg)$degree <- degree(ingT_deg) # assignment

# plot(ingT_deg, 
#     vertex.label.cex = .6, 
#     vertex.label.color = "black", 
#     vertex.size = V(ingTypeEL_Dig)$degree, 
#     vertex.label.cex = .2) # sized by degree

@@

view(betweenness(ingTypeEL_Dig, directed = FALSE))

# view(betweenness(ingTypeEL_Dig, directed = TRUE))
              # Yeilds 0

ingTypeEL_Between <- ingTypeEL_Dig
# We can again assign the output of betweenness() to a variable in the network and size the nodes according to it.

V(ingTypeEL_Between)$betweenness <- 
  betweenness(ingTypeEL_Between, 
              directed = F)

# plot(ingTypeEL_Between, 
#     vertex.label.cex = .6, 
#     vertex.label.color = "black", 
#     vertex.size = V(ingTypeEL_Between)$betweenness) # sized by betweenness

@@

view(evcent(ingTypeEL_Dig)$vector)
# view(evcent(smugEL_ig)$vector)
ingTyp_Vector <- ingTypeEL_Dig
V(ingTyp_Vector)$eigenvector <- evcent(ingTyp_Vector)$vector

@@

ingTyp_Cent <- ingTypeEL_Dig

>>>>

---
title: "Week 3 Notebook"
output: html_notebook

# Network Info
## Smuggler's Cove Index- NETWORK: EDGELIST: Features & Descriptions

@@

            # list the names:  edge attributes
  network::list.edge.attributes(smugEL_sn)
      #   summarize: numeric network attribute
  summary(smugEL_sn%v%"dnkNam")

@@

@@

@@

    # Generate a list the names of vertex attributes
  view(V(smugEL_ig)$name)

    # access edge attribute  names
  view(igraph::vertex_attr_names(smugEL_ig))

    # Generate a list the names of edge attributes  
  view(igraph::edge_attr_names(smugEL_ig))

## Network Attributes
 summary(smugEL_ig)
          # Vertices =  
          # Directed = 
          # hyper = 
          # loops = 
          # multiple = 
          # bipartite = 
          # total edges = 
                # missing edges =
                # non-missing edges = 
          # Vertex attribute names:
          # Edge Attribute Names: 
```

```{r IGRAPH: DIRECTED Network: ingTyp, echo=TRUE}

      # class of an object
    class(ingTypEL_Dig)

@@

      # Count Vertices
    vcount(ingTypEL_Dig)

      # Count Edges
    ecount(ingTypEL_Dig)

## Network Mode
    # Bipartitie / Unimodal
    is_bipartite(ingTypEL_Dig)

## Edges    
    # Un/directed 
    is_directed(ingTypEL_Dig)

    # weighted or unweighted?
    is_weighted(ingTypEL_Dig)

    # edge weights
    E(ingTypEL_Dig)$weight

@@

    # Generate a list the names of vertex attributes
  V(ingTypEL_Dig)$name

    # access edge attribute  names
  igraph::vertex_attr_names(ingTypEL_Dig)

    # Generate a list the names of edge attributes  
  igraph::edge_attr_names(ingTypEL_Dig)

## Network Attributes
 summary(ingTypEL_Dig)
          # Vertices =  
          # Directed = 
          # hyper = 
          # loops = 
          # multiple = 
          # bipartite = 
          # total edges = 
                # missing edges =
                # non-missing edges = 
          # Vertex attribute names:
          # Edge Attribute Names:

@@

#     find average shortest path for network
    average.path.length(ingTypEL_Dig,directed=T)

@@

#     isolating the distances between specific node
    all_shortest_paths(ingTypEL_Dig,
                       "Strozzi","Tornabuoni", 
                       weights=NA)$res
## Note: manually tell igraph to ignore edge weights.
```

@@

#     Calculate distances between two nodes
    distances(ingTypEL_Dig,"Bischeri","Ridolfi")

#     Calculate distance between two nodes using unweighted edges
    distances(ingTypEL_Dig,"Bischeri", "Castellani",weights=NA)

@@

    # Number of components
 igraph::components(smugEL_ig)$no 
      ##**NOTE:** $no prevents a list of components from being printed

@@

```{r Statnet: , echo=TRUE}
      # Basic R: get summary statistics for node attributes
 summary(nodes_ingTyps)
      # provides range, mean, median and standard deviation of a vector.

@@

smug_attributes <- smugCovIdx
smug_edj <- smugCovIdx[2:3]
View(smug_edj)
#V1: dnkIng
#V2: ingTyp

@@

V(smugNet)$dnkIng <- 
  smug_attributes$dnkIng[match(V(smugNet)$name, smug_attributes$ingTyp)]

V(smugNet)$dnkNam <- 
  smug_attributes$dnkNam[
    match(V(
      smugNet)$name,
      smug_attributes$dnkIng)
    ]
summary(smugNet)

@@

V(smugNet)$ingTyp <- 
  smug_attributes$ingTyp[match(V(smugNet)$name, smug_attributes$ingTyp)]

V(smugNet)$dnkNam <-
  smug_attributes$dnkNam[match(V(smugNet)$name, smug_attributes$dnkNam)]

@@

```{r iGraph: Undirected: DEGREE CENTRALITY: ingTyp, echo=TRUE}
# graph the SMUG network

smug_attributes <- smugCovIdx
smug_edj <- smugCovIdx[2:3]
View(smug_edj)
#V1: dnkIng
#V2: ingTyp

@@

V(smugNet)$dnkNam <-
  smug_attributes$dnkNam[match(V(smugNet)$name, smug_attributes$dnkIng)]

@@

 # differentiate nodes based on their attributes
 gplot(smugEL_sn, 
       displaylabels=TRUE,
       label.cex=.75,
       label.col="black",
       vertex.col=att$dnkNam
 )

@@

graph.density(ingTypEL_Dig)
    ## Density captures, at the macro-level, the proportion of dyads that are present over the possible total number of dyads in the network

@@

ingTypEL_DigNod <- vcount(ingTypEL_Dig)

ingTypEL_DigDen <- graph.density(ingTypEL_Dig)

@@

```{r iGraph: Directed: Degree Centrality}
# Degree centrality is calculated using the degree function in R. It returns how many edges each node has.
view(degree(ingTypEL_Dig))

ingT_deg <- ingTypEL_Dig
V(ingT_deg)$degree <- degree(ingT_deg) # assignment

# plot(ingT_deg, 
#     vertex.label.cex = .6, 
#     vertex.label.color = "black", 
#     vertex.size = V(ingTypEL_Dig)$degree, 
#     vertex.label.cex = .2) # sized by degree

@@

view(betweenness(ingTypEL_Dig, directed = FALSE))

# view(betweenness(ingTypEL_Dig, directed = TRUE))
              # Yeilds 0

ingTypEL_Between <- ingTypEL_Dig
# We can again assign the output of betweenness() to a variable in the network and size the nodes according to it.

V(ingTypEL_Between)$betweenness <- 
  betweenness(ingTypEL_Between, 
              directed = F)

# plot(ingTypEL_Between, 
#     vertex.label.cex = .6, 
#     vertex.label.color = "black", 
#     vertex.size = V(ingTypEL_Between)$betweenness) # sized by betweenness

@@

view(evcent(ingTypEL_Dig)$vector)
# view(evcent(smugEL_ig)$vector)
ingTyp_Vector <- ingTypEL_Dig
V(ingTyp_Vector)$eigenvector <- evcent(ingTyp_Vector)$vector

@@

ingTyp_Cent <- ingTypEL_Dig
